\vspace{-2cm} Government websites as data: \\ A methodological pipeline with application to the websites of municipalities in the United States

### 2019-05-06 paper/manuscript_paNote.tex

\section{Introduction}
Local governments convey voluminous information about all aspects of their policymaking, policy implementation, and public deliberation, via their official websites. The vital role of official websites in connecting the government and the governed has motivated a wave of research on the contents of government websites \citep[e.g.,][]{grimmelikhuijsen2010transparency,wang2005evaluating,osman2014cobra,Eschenfelder2002}. Despite the potential for automated scraping of website contents, the conventional approach to data collection in projects focused on government websites involves manual content extraction from each website in the dataset. Though accurate, the manual approach to data collection is costly, and cannot be scaled to capture even a fraction of the volume of content available on government websites. In this paper, we present a methodological pipeline that can be used to automatically scrape government websites in order to build datasets that can be used for text analysis---describing challenges in data collection and processing, as well as the solutions we adopt. We provide an illustrative application in which we explore the ways in which the textual contents on city government websites in six American states (IN, LA, NY, WA, CA and TX) correlate with the partisanship of the city mayor.

@@

 A substantial body of research has found that the partisanship of the mayor affects city governance along multiple dimensions of spending and policy attention \citep{gerber2011mayors,de2016mayoral,einstein2016mayors,marion2013mayor}. Official city websites allow mayors to present their views and policy priorities to the public. In local politics, where campaign funds are low, this lends incumbents a crucial advantage in becoming more well-known among their constituencies \citep{stanyer2008elected}. Local government websites are frequently visited by the public \citep{thomas2003new}. City websites can be used to communicate the stance of a mayor on social or economic programs. In the current paper, we evaluate the feasibility and performance of text analysis of city government websites in discerning differences in city governments based on mayoral partisanship.

@@

The existing research that uses scraped websites provides an indication of the theoretical value of empirical analysis of web contents. The most pertinent literature to our research is the e-governance literature, which focuses on the online presence of governments from a usability and public service point of view. Research in this category evaluates government websites in terms of accessibility, ease-of-use, and function \citep[e.g., ][]{Urban2002,mcnutt2010virtual,Armstrong2011,Feeney2017}. As an example, \cite{grimmelikhuijsen2012developing} study local government websites with the goal of uncovering how they aid the goal of transparency. To this end, they analyze a set of Dutch municipalities in which air quality had deteriorated. The authors test whether local governments provide citizens with information about potential complications and solutions associated with this issue.

The websites of politicians and their parties have also been the object of research. Researchers have found that in order to identify the constituencies, motives, and modes of communication among these actors, their websites can be very illuminating sources of information \citep{Druckman2009,Druckman2010,Cryer2017,Esterling2011,Esterling2011a,Norris2003,Therriault2010}. For example, \citep{Druckman2010} analyze the issues engaged on websites for candidates in U.S. Congressional elections, and find that candidates strategically engage just a few issues based on the priorities in their districts and the characteristics of their opponents.

@@

For data availability reasons, we focus our analysis of municipal websites on six states---Indiana, Louisiana, New York, Washington, California, and Texas. [{\bf Markus, when were the websites scarped---need to indicate year and/or month}]. The selection of states and cities is largely dictated by the presence of partisan mayors and availability of the relevant data. Municipal elections in Indiana and Louisiana are partisan across the board,\footnote{In Indiana, this includes only cities - incorporated municipalities with at least 2,000 inhabitants - as opposed to towns.} so our sample is primarily focused on these two states. For Indiana and Louisiana, all cities with a website are included, resulting in a considerably larger sample than for the other four states. New York and Washington do not have nominally partisan elections, but for a subset of cities, partisanship can be determined through contribution data (see below for more detail on how we use this data). Finally, California and Texas contain a number of large cities, whose mayors are sufficiently well-known for their partisanship to be available. These six states also provide us with a sample that is well-balanced on a number of theoretically important indicators. One, each of the four geographic regions [{\bf Markus, are these census regions?}] are represented with at least one state. Two, we have a fairly well-balanced sample with respect to the urban/rural cleavage, as both major cities and less densely populated areas are covered. Furthermore, the sample is politically balanced---we have three blue states (CA, WA, NY) and three red states (TX, IN, LA).  The partisan breakdown of city websites is depicted in Table \ref{tab:cityparties}. The dataset has a relative party balance, with variation in each state. Details on the sources and methods of raw data collection can be found in the Appendix.

@@

One of the more subtle aspects of local government is the presence of different types of government structures. Between council-manager governments and mayor-council governments \citep{morgan1992policy}---either in the weak or strong mayor variant \citep{desantis2002city}---there is a certain degree of variance in where a city's executive authority lies. Unfortunately, we do not have access to information about the type of governments across the breadth of our dataset. Given the prominent place that mayors tend to have on their cities' websites, we feel that any bias arising from this nuance should be minor. \cite{gerber2011mayors}, whose model is somewhat comparable to ours in the sense that they also test the effect of mayoral partisanship on city policy priorities.

@@

The results are shown in Table \ref{tabSTMtopwords60}. The rows of the table are sorted so that the most Republican topics (marked by a deeper red color) appear at the top, and the most Democratic ones (blue) at the bottom. The topics that are entirely white have 90\% credible intervals on the effect of the mayoral party that include zero.

Many of the topics associated with Democrats fit with what we understand to be national party priorities. Topic 21, on affordable housing, clearly resonates with the Democratic party's appeal to low-income voters. Similarly, employee rights are represented in Topic 47. Democrats also exhibit a strong preference for words related to public finances, such as Topic 32 ('budget', 'revenue', 'expenditure') as well as Topic 19 ('debt', 'bond', 'financial'). We suspect that the association of Democratic mayors with finance-related terms is indicative of a greater willingness to emphasize the city's efforts to raise and spend money. This finding is consistent with \citep{Einstein2015}, who show that Democratic mayors tend to favor greater spending. A second, consistent Democratic focus appears to be law enforcement: The most Democratic topic, 55 ('robbery', 'homicide', 'sergeant') (a comparable topic is also the most Democratic topic in the model with 120 topics in tables \ref{tabSTMtopwords120_1} and \ref{tabSTMtopwords120_2} of the Appendix) depicts Democrats' complicated relationship with law enforcement. On the one hand, Democratic partisans have a more negative perception of the police, rating it considerably more negatively on the appropriate use of force and the equal treatment of minorities \citep{Brown2017}. On the other hand, the literature has also shown that cities with a higher Democratic vote share spend more on the police, even after controlling for crime \citep{Einstein2015}. Finally, Democrats also focus more on the deliberative process of policymaking, as topics 31 ('agenda', 'committee'), 34 ( `comment', `draft', `feedback'), 48 (`absent', `aye', `nay'), and 37 ('audit', 'procedure', 'oversight') attest to.  This openness regarding the policy process on behalf of cities with Democratic mayors fits with the findings of \citet{grimmelikhuijsen2012developing}, which are that left-wing local governments exhibit greater transparency via website content.

@@

The partisanship of the mayor of each city is coded in different ways, depending on the state. For Indiana, where elections are nominally partisan, this information is accessible through the state government's website\footnote{\url{http://www.in.gov/apps/sos/election/general/general2015?page=office&countyID=1&officeID=32&districtID=-1&candidate=}}. For Louisiana, we received data on the outcomes of mayoral elections from the Local Elections in America Project (LEAP) \citep{marschall2013local}. For the other states, where mayoral elections are not nominally partisan (but the partisanship of the mayor is still well-known), we employed different means: For New York and Washington, we searched the state campaign finance websites, and coded the parties of the candidates based on the party committees from which they received donations. For California and Texas, where our data consists of highly populated cities, partisanship information was acquired from Ballotpedia\footnote{\url{https://ballotpedia.org/List_of_current_mayors_of_the_top_100_cities_in_the_United_States}}. Finally, we also scraped mayoral partisanship from the cities' Wikipedia pages. When compared to the other data sources above, (and manual searches in case of conflicts) Wikipedia proved to be very reliable and added additional cases to our dataset even for Indiana and Louisiana. Generally speaking, we found data scraped from Wikipedia, aided by manual corrections in case of missing or conflicting data, to be more reliable than data from governmental sources.

>>>>

\section{Introduction}
Local governments convey voluminous information about all aspects of their policymaking, policy implementation, and public deliberation, via their official websites. The vital role of official websites in connecting the government and the governed has motivated a wave of research on the contents of government websites \citep[e.g.,][]{grimmelikhuijsen2010transparency,wang2005evaluating,osman2014cobra,Eschenfelder2002}. The conventional approach to data collection in projects focused on government websites involves manual content extraction from each website in the dataset. Though accurate, the manual approach to data collection is costly for large-scale analysis. We present a methodological pipeline that can be used to automatically scrape government websites in order to build datasets that can be used for text analysis---describing challenges in data collection and processing, as well as the solutions we adopt. We provide an illustrative application in which we explore the ways in which the textual contents on city government websites in six American states correlate with the partisanship of the city mayor.

@@

 A substantial body of research has found that the partisanship of the mayor affects city governance along multiple dimensions of spending and policy attention \citep{gerber2011mayors,de2016mayoral,einstein2016mayors,marion2013mayor}. Official city websites allow mayors to present their views and policy priorities to the public. In local politics, where campaign funds are low, this lends incumbents a crucial advantage in becoming more well-known among their constituencies \citep{stanyer2008elected}. Local government websites are frequently visited by the public \citep{thomas2003new}. City websites can be used to communicate the stance of a mayor on social or economic programs.

@@

The existing research that uses scraped websites provides an indication of the theoretical value of empirical analysis of web contents. Research `e-governance' evaluates government websites in terms of accessibility, ease-of-use, and function \citep[e.g., ][]{Urban2002,mcnutt2010virtual,Armstrong2011,Feeney2017}. As an example, \cite{grimmelikhuijsen2012developing} study local government websites of Dutch municipalities to measure government transparency regarding air quality in the municipalities. The websites of politicians and their parties have also been the object of research \citep{Druckman2009,Druckman2010,Cryer2017,Esterling2011,Esterling2011a,Norris2003,Therriault2010}. For example, \citep{Druckman2010} analyze the issues engaged on websites for candidates in U.S. Congressional elections, and find that candidates strategically engage just a few issues based on the priorities in their districts and the characteristics of their opponents.

@@

For data availability reasons, we focus our analysis of municipal websites on six states---Indiana, Louisiana, New York, Washington, California, and Texas. [{\bf Markus, when were the websites scarped---need to indicate year and/or month}]. The selection of states and cities is largely dictated by the presence of partisan mayors and availability of the relevant data. Municipal elections in Indiana and Louisiana are partisan across the board, so our sample is primarily focused on these two states. For Indiana and Louisiana, all cities with a website are included, resulting in a considerably larger sample than for the other four states. New York and Washington do not have nominally partisan elections, but for a subset of cities, partisanship can be determined through contribution data (see appendix for more detail). California and Texas contain a number of large cities whose mayors are sufficiently well-known for their partisanship to be available. Our sample is well-balanced on a number of theoretically important dimension. One, each of the four geographic regions [{\bf Markus, are these census regions?}] are represented with at least one state. Two, we have a fairly well-balanced sample with respect to the urban/rural cleavage. Furthermore, the sample is politically balanced---we have three blue states, and three red states.  The partisan breakdown of city websites is depicted in Table \ref{tab:cityparties}. Details on the sources and methods of raw data collection can be found in the Appendix.

@@

One of the more subtle aspects of local government is the presence of different types of government structures. Between council-manager governments and mayor-council governments \citep{morgan1992policy}---either in the weak or strong mayor variant \citep{desantis2002city}---there is variance in where a city's executive authority lies. We do not have access to information about the type of governments across the breadth of our dataset. Given the prominent place that mayors tend to have on their cities' websites, we feel that any bias arising from this nuance should be minor. \cite{gerber2011mayors}, whose model is somewhat comparable to ours in the sense that they also test the effect of mayoral partisanship on city policy priorities.

@@

The results are shown in Table \ref{tabSTMtopwords60}. Many of the topics associated with Democrats fit with what we understand to be national party priorities. Topic 21, on affordable housing, clearly resonates with the Democratic party's appeal to low-income voters. Similarly, employee rights are represented in Topic 47. Democrats also exhibit a strong preference for words related to public finances, such as Topic 32 ('budget', 'revenue', 'expenditure') as well as Topic 19 ('debt', 'bond', 'financial'). We suspect that the association of Democratic mayors with finance-related terms is indicative of a greater willingness to emphasize the city's efforts to raise and spend money. This finding is consistent with \citep{Einstein2015}, who show that Democratic mayors tend to favor greater spending. A second, consistent Democratic focus appears to be law enforcement: The most Democratic topic, 55 ('robbery', 'homicide', 'sergeant') (a comparable topic is also the most Democratic topic in the model with 120 topics in tables \ref{tabSTMtopwords120_1} and \ref{tabSTMtopwords120_2} of the Appendix) depicts Democrats' complicated relationship with law enforcement. On the one hand, Democratic partisans have a more negative perception of the police, rating it considerably more negatively on the appropriate use of force and the equal treatment of minorities \citep{Brown2017}. On the other hand, the literature has also shown that cities with a higher Democratic vote share spend more on the police, even after controlling for crime \citep{Einstein2015}. Finally, Democrats also focus more on the deliberative process of policymaking, as topics 31 ('agenda', 'committee'), 34 ( `comment', `draft', `feedback'), 48 (`absent', `aye', `nay'), and 37 ('audit', 'procedure', 'oversight') attest to.  This openness regarding the policy process on behalf of cities with Democratic mayors fits with the findings of \citet{grimmelikhuijsen2012developing}, which are that left-wing local governments exhibit greater transparency via website content.

@@

The partisanship of the mayor of each city is coded in different ways, depending on the state. For Indiana, where elections are nominally partisan, this information is accessible through the state government's website\footnote{\url{http://www.in.gov/apps/sos/election/general/general2015?page=office&countyID=1&officeID=32&districtID=-1&candidate=}}. For Louisiana, we received data on the outcomes of mayoral elections from the Local Elections in America Project (LEAP) \citep{marschall2013local}. For the other states, where mayoral elections are not nominally partisan (but the partisanship of the mayor is still well-known), we employed different means: For New York and Washington, we searched the state campaign finance websites, and coded the parties of the candidates based on the party committees from which they received donations. For California and Texas, where our data consists of highly populated cities, partisanship information was acquired from Ballotpedia\footnote{\url{https://ballotpedia.org/List_of_current_mayors_of_the_top_100_cities_in_the_United_States}}. Finally, we also scraped mayoral partisanship from the cities' Wikipedia pages. When compared to the other data sources above, (and manual searches in case of conflicts) Wikipedia proved to be very reliable and added additional cases to our dataset even for Indiana and Louisiana. Generally speaking, we found data scraped from Wikipedia, aided by manual corrections in case of missing or conflicting data, to be more reliable than data from governmental sources.\footnote{In Indiana, the data includes only cities - incorporated municipalities with at least 2,000 inhabitants - as opposed to towns.}
